---
title: "Lead IQ Data Reprot"
author: "Jess Wild"
date: "September 12, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Package Instalation and Data Import
```{r packages, echo=FALSE}
#instlling needed packages
require(ggplot2)
require(kableExtra)
require(table1)
```

```{r data}
#import the raw data 
dat <- read.csv("C:/Users/jessi/Desktop/CU Denver Biostat PhD/6621/HW/HW 2/lead-iq-01.csv")

```

## Data Description

This data set includes information on the IQ of 124 people and a categorical variable describing how close they lived to a lead smelter.  THe purpose of this dataset is to investigat the relationship between lead exposure and measured IQ level.

### Variable 1: Smelter

If a subject lives within one mile of a lead smelter, they are classified as 'Near', otherwise they are classified as 'Far'.  There were `r nrow(subset(dat,dat$Smelter=='Near'))` observations classified as 'Near' and `r nrow(subset(dat,dat$Smelter=='Far'))` observations classified as 'Far'.

### Variable 2: IQ

Using the Wechsler Intelligence Scale for Children, each subject's IQ score was recorded.  The mean IQ score overall was `r mean(dat$IQ)`. Thea mean IQ score for those calssified as near to a smelter was `r mean(subset(dat,dat$Smelter=='Near')$IQ)`, and the mean for those classified as far from a smelter was `r mean(subset(dat,dat$Smelter=='Far')$IQ)`.

### Data Tables
#### Data

The full data set is:

```{r data table}
knitr::kable(dat, caption = "TABLE 1")

```
 
#### Descriptive Statistics

Some descriptive statistics for this data are:

```{r descriptive table}
table1(~IQ+Smelter|Smelter, data = dat)

```

## Data Visualization

To visualize the data, we can generate a boxplot stratified by proximity to a smelter:


```{r boxplot, echo=FALSE}
ggplot(dat, aes(x=Smelter, y=IQ, fill=Smelter)) +
  geom_boxplot()
```


We can also generate a histogram of the data:

```{r histogram, echo=FALSE}
ggplot(dat, aes(x=IQ, fill=Smelter, color=Smelter)) +
  geom_histogram(position="identity", alpha=0.5)
```